The Leaky Semicolon

### 2021-10-28 paper/mca2.tex
removing todos. Integrated consistency and causal strengtheining.

\armeight{} allows the execution because the read of $x$ is internal to the
thread.  This aspect of \armeight{} semantics is difficult to model locally.
To capture this, we found it necessary to drop \ref{pom-rf-le} and relax
\ref{seq-le-delays}, %from
% \refdef{def:pwt:mca1},
adding local constraints on $\rrfx$ to $\sPAR{}{}$, $\sSEMI{}{}$ and
$\sIF{}{}{}$.
Rather than ensuring that there is no
\emph{global} blocker for a sequentially fulfilled read \eqref{pom-rf-le}, we
require only that there is no \emph{thread-local} blocker
\eqref{seq-le-rf-rf}.
For \PwTmca{2}, internal reads don't necessarily contribute to order, and
thus the above execution is allowed.

\begin{definition}
  \label{def:pwt:mca2}
  The definition of \PwTmca{2} is derived from that of \PwTmca{1} by
  removing \ref{pom-rf-le} and adding the following:

  \begin{enumerate}[topsep=0pt]
  \item[{\labeltext[\textsc{p}6a]{(\textsc{p}6a)}{par-le-rf1}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\bEv\xrfx\aEv$ then $\bEv\lt\aEv$,
  \item[{\labeltext[\textsc{p}6b]{(\textsc{p}6b)}{par-le-rf2}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\aEv\xrfx\bEv$ then $\aEv\lt\bEv$,
  \item[{\labeltext[\textsc{s}6a$'$]{(\textsc{s}6a$'$)}{seq-le-delays-rf}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\labeling_1(\bEv) \rdelays \labeling_2(\aEv)$ 
    then either $\bEv\xrfx\aEv$ or $\bEv\le\aEv$,
  \item[{\labeltext[\textsc{s}6b]{(\textsc{s}6b)}{seq-le-rf-rf}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\labeling_1(\bEv) \rblocks \labeling_2(\aEv)$
    then $\cEv\xrfx\aEv$ implies $\bEv\le\cEv$.
    % \\
    % if $\labeling_1(\cEv) \rblocks \labeling_2(\aEv)$ and $\bEv\xrfx\aEv$
    % then $\cEv\le\bEv$.
  % \item[{\labeltext[\textsc{i}6a]{(\textsc{i}6a)}{if-le-rf}}]
  %   ${\rrfx}\subseteq {\rrfx_1}\cup {\rrfx_2}$
  \end{enumerate}
  % The definition of \PwTmca{2} is a \PwT{} (\refdef{def:pomset}) equipped with an injective
  % relation $\rrfx$ that satisfies requirements \ref{pom-rf-match} and
  % \ref{pom-rf-block} of \refdef{def:pwt:mca1}.

@@

  %\input{fig-mca2.tex}
\end{definition}
\ref{par-le-rf1} and \ref{par-le-rf2} ensure that $\bEv\xrfx\aEv$ implies
$\bEv\lt\aEv$ when the actions come from different threads.  However, we may
have $\bEv\xrfx\aEv$ and $\aEv\lt\bEv$ within a thread, as between
$\DWP{x}{2}$ to $\DRP[\mACQ]{x}{2}$ in \ref{internal-acq}.  To allow this,
\ref{seq-le-delays-rf} relaxes \ref{seq-le-delays} from \PwTmca{1}.
\ref{seq-le-rf-rf} is then necessary to stop stale reads within a thread.
For example, it prevents a read of $1$ in
\begin{math}
  \PW{x}{1}
  \SEMI
  \PW{x}{2}
  \SEMI
  \PR{x}{r}.
\end{math}
\todo{Is s6b necessary?}

@@

With the weakening of \ref{seq-le-delays}, we must be careful not to allow
spurious pairs to be added to the $\rrfx$ relation.  For example, %\ref{if-rf-project} ensures that
\begin{math}
  \frf{\semrr{\IF{b}\THEN\PR{x}{r}\PAR\PW{x}{1}\ELSE\PR{x}{r}\SEMI\PW{x}{1}\FI}}
\end{math}
should not include 
\begin{math}
  \smash{\hbox{\begin{tikzinlinesmall}[node distance=1.5em]
        \event{a}{\DR{x}{1}}{}
        \event{b}{\DW{x}{1}}{right=of a}
        \rfint[out=165,in=15]{b}{a}
        \wki{a}{b}
      \end{tikzinlinesmall}}}
\end{math}, taking $\rrfx$ from the left and $\lt$ from the right.
The use of $\rextends{}{}$ ensures this.

% \PwTmca{2} does not enforce \ref{pom-rf-le}: $\bEv\xrfx\aEv$ may not imply
% $\bEv\le\aEv$ when $\bEv$ and $\aEv$ come from different sides of a
% sequential composition.  This means that
As a consequence of dropping \ref{pom-rf-le}, sequential $\rrfx$ must be validated during
pomset construction, rather than post-hoc.  In \textsection\ref{sec:c11}, we
show how to construct program order ($\rpox$) for complete pomsets using
phantom events ($\fmrg{}$).  Using this construction, the following lemma
gives a post-hoc verification technique for $\rrfx$.  Let $\fmrginv{}$ be the
inverse of $\fmrg{}$.
\begin{lemma}
  \label{lem:mca2}
  If $\aPS\in\frf{\semmca{2}{\aCmd}}$ is complete, then
  for every $\bEv\xrfx\aEv$ either
  \begin{itemize}
  \item external fulfillment:
    $\bEv\lt\aEv$ and if $\labeling(\cEv) \rblocks\labeling(\aEv)$ then either $\cEv\le\bEv$ or $\aEv\le\cEv$, or
  \item internal fulfillment:
    $(\exists\bEv'\in\fmrginv{\bEv})$
    $(\exists\aEv'\in\fmrginv{\aEv})$
    $\bEv'\xpox\aEv'$ and $(\not\exists\cEv)$
    $\labelingForm(\cEv)$ is a tautology and
    $\labeling(\cEv) \rblocks \labeling(\aEv)$ and $\bEv'\xpox\cEv\xpox\aEv'$.
  \end{itemize}
\end{lemma}
These mimic the \emph{external consistency} requirements of \armeight{}
\cite{armed}.

>>>>

\armeight{} allows the execution because the read of $x$ is internal to the
thread.  This aspect of \armeight{} semantics is difficult to model locally.
To capture this, we found it necessary to drop \ref{pom-rf-le} and relax
\ref{seq-le-delays}, %from
% \refdef{def:pwt:mca1},
adding local constraints on $\rrfx$ to $\sPAR{}{}$, $\sSEMI{}{}$ and
$\sIF{}{}{}$.
% Rather than ensuring that there is no
% \emph{global} blocker for a sequentially fulfilled read \eqref{pom-rf-le}, we
% require only that there is no \emph{thread-local} blocker
% \eqref{seq-le-delays-rf}.
% For \PwTmca{2}, internal reads don't necessarily contribute to order, and
% thus the above execution is allowed.
(For parallelism, we explicitly specify the domain of $\bEv$ and $\aEv$ in
\ref{seq-le-delays-rf}.)

\begin{definition}
  \label{def:pwt:mca2}
  The definition of \PwTmca{2} is derived from that of \PwTmca{1} by
  removing \ref{pom-rf-le} and \ref{seq-le-delays} and adding the following:

  \begin{enumerate}[topsep=0pt]
  \item[{\labeltext[\textsc{p}6a]{(\textsc{p}6a)}{par-le-rf1}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\bEv\xrfx\aEv$ then $\bEv\lt\aEv$,
  \item[{\labeltext[\textsc{p}6b]{(\textsc{p}6b)}{par-le-rf2}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\aEv\xrfx\bEv$ then $\aEv\lt\bEv$,
  \item[{\labeltext[\textsc{s}6a$'$]{(\textsc{s}6a$'$)}{seq-le-delays-rf}}]
    if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\labeling_1(\bEv) \rdelays \labeling_2(\aEv)$ 
    then either $\bEv\xrfx\aEv$ or $\bEv\le\aEv$,
  % \item[{\labeltext[\textsc{s}6b]{(\textsc{s}6b)}{seq-le-rf-rf}}]
  %   if $\bEv\in\aEvs_1$, $\aEv\in\aEvs_2$ and $\labeling_1(\bEv) \rblocks \labeling_2(\aEv)$
  %   then $\cEv\xrfx\aEv$ implies $\bEv\le\cEv$.
    % \\
    % if $\labeling_1(\cEv) \rblocks \labeling_2(\aEv)$ and $\bEv\xrfx\aEv$
    % then $\cEv\le\bEv$.
  % \item[{\labeltext[\textsc{i}6a]{(\textsc{i}6a)}{if-le-rf}}]
  %   ${\rrfx}\subseteq {\rrfx_1}\cup {\rrfx_2}$
  \end{enumerate}
  % The definition of \PwTmca{2} is a \PwT{} (\refdef{def:pomset}) equipped with an injective
  % relation $\rrfx$ that satisfies requirements \ref{pom-rf-match} and
  % \ref{pom-rf-block} of \refdef{def:pwt:mca1}.

@@

  %\input{fig-mca2.tex}
\end{definition}
\ref{par-le-rf1} and \ref{par-le-rf2} ensure that $\bEv\xrfx\aEv$ implies
$\bEv\lt\aEv$ when the actions come from different threads.  However, we may
have $\bEv\xrfx\aEv$ and $\aEv\lt\bEv$ within a thread, as between
$\DWP{x}{2}$ to $\DRP[\mACQ]{x}{2}$ in \ref{internal-acq}, thus allowing this
execution.
% To accomplish this, \ref{seq-le-delays-rf} relaxes
% \ref{seq-le-delays} from \PwTmca{1}.
\ref {pom-rf-block} and \ref{seq-le-delays-rf} are sufficient to stop stale
reads within a thread.  For example, it prevents a read of $1$ in
\begin{math}
  \PW{x}{1}
  \SEMI
  \PW{x}{2}
  \SEMI
  \PR{x}{r}.
\end{math}

@@

With the weakening of \ref{seq-le-delays}, we must be careful not to allow
spurious pairs to be added to the $\rrfx$ relation.  For example, %\ref{if-rf-project} ensures that
\begin{math}
  \frf{\semrr{\IF{b}\THEN\PR{x}{r}\PAR\PW{x}{1}\ELSE\PR{x}{r}\SEMI\PW{x}{1}\FI}}
\end{math}
should not include 
\begin{math}
  \smash{\hbox{\begin{tikzinlinesmall}[node distance=1.5em]
        \event{a}{\DR{x}{1}}{}
        \event{b}{\DW{x}{1}}{right=of a}
        \rfint[out=-165,in=-15]{b}{a}
        \wki[out=15,in=165]{a}{b}
      \end{tikzinlinesmall}}}
\end{math}, taking $\rrfx$ from the left and $\lt$ from the right.
The use of ``$\rextends{}{}$'' in \ref{if-le} and \ref{if-rf} ensures this.

% \PwTmca{2} does not enforce \ref{pom-rf-le}: $\bEv\xrfx\aEv$ may not imply
% $\bEv\le\aEv$ when $\bEv$ and $\aEv$ come from different sides of a
% sequential composition.  This means that
As a consequence of dropping \ref{pom-rf-le}, sequential $\rrfx$ must be validated during
pomset construction, rather than post-hoc.  In \textsection\ref{sec:c11}, we
show how to construct program order ($\rpox$) for complete pomsets using
phantom events ($\fmrg{}$).  Using this construction, the following lemma
gives a post-hoc verification technique for $\rrfx$.  Let $\fmrginv{}$ be the
inverse of $\fmrg{}$.
\begin{lemma}
  \label{lem:mca2}
  If $\aPS\in\frf{\semmca{2}{\aCmd}}$ is complete, then
  for every $\bEv\xrfx\aEv$ either
  \begin{itemize}
  \item external fulfillment:\;\,
    $\bEv\lt\aEv$ and if $\labeling(\cEv) \rblocks\labeling(\aEv)$ then either $\cEv\le\bEv$ or $\aEv\le\cEv$, or
  \item internal fulfillment:
    \begin{math}
      \begin{array}[t]{l}
        (\exists\bEv'\in\fmrginv{\bEv})
        (\exists\aEv'\in\fmrginv{\aEv})
        \\ 
        \bEv'\xpox\aEv' \textand (\not\exists\cEv)\;
        % \labelingForm(\cEv) is a tautology \textand
        \labeling(\cEv) \rblocks \labeling(\aEv) \textand
        \bEv'\xpox\cEv\xpox\aEv'.        
      \end{array}
    \end{math}
  \end{itemize}
\end{lemma}
These mimic the \emph{external consistency} requirements of \armeight{}
\cite{armed}.

### 2021-10-28 paper/outro.tex
Acknowledgements: Anton
done with section 2
removing todos. Integrated consistency and causal strengtheining.

\section{Conclusions} %Future Work}
\label{sec:outro}

This paper is the first to present a direct denotational semantics for
sequential composition in a relaxed memory model that can be
efficiently compiled to modern CPUs. We extract from this model a semantic
dependency relation and use it to build \PwTc, a solution to the
Out-of-Thin-Air problem in \cXI, and \PwTmca{}, a model for Java-like
languages.
% There is, as usual, more research to be done.

We have not treated loops in this model, though we expect that the usual
approach of showing continuity for all the semantic operations with respect
to set inclusion would go through. \citet{DBLP:conf/esop/PaviottiCPWOB20} use
step-indexing to account for loops; a similar approach could be applied here.

@@

\PwTmca{1} is a simpler model than \PwTmca{2}, but requires fences on
acquiring reads for \armeight{}.  It would be illuminating to find out what
the performance penalty is for these fences.

% In a similar vein,
% \PwT{} does not validate read introduction, whereas speculative operational
% semantics do.  Recent work shows a tension between read introduction and
% compositional reasoning for temporal safety properties (see
% \textsection\ref{sec:related}).  Nonetheless, read introduction is
% ubiquitous in some compilers.  It would be interesting to know if there is a
% performance penalty for banning read introduction.

@@

@@

Riely was supported by the National Science Foundation under
grant No.~CCR-1617175.

@@

\todo{Add support for Anton\&Ilya.}

>>>>

\section{Limitations and Future Work}
\label{sec:outro}

This paper is the first to present a direct denotational semantics for
sequential composition in a relaxed memory model that can be efficiently
compiled to modern CPUs. We extract from this model a semantic dependency
relation and use it to build \PwTc, a solution to the Out-of-Thin-Air problem
in \cXI, and \PwTmca{}, a model intended for safe languages such as Java and
Javascript.  Our work has several limitations, providing opportunities for
future work.

We have mechanized some proofs, but not all.  In particular, we have only a
pen-and-paper proof showing that \PwTmca{} supports optimal lowering to
\armeight{}.  The same is true for local data race-freedom (\ldrfsc{}).
Additionally, our proof sketch for \ldrfsc{} elides \RMW{}s, which have
caused complications in other models \cite{promising-ldrf}.

We have not treated loops, although we expect that the usual approach of
showing continuity for all the semantic operations with respect to set
inclusion would go through. \citet{DBLP:conf/esop/PaviottiCPWOB20} use
step-indexing to account for loops; perhaps this approach could be adapted.

@@

\PwTmca{1} is a simpler model than \PwTmca{2}, but requires fences on
acquiring reads for \armeight{}.  It would be illuminating to find out what
the performance penalty is for these fences.  In a similar vein, it would be
interesting to know if there is a performance penalty for banning
read-introduction, which is not valid for \PwT{}.

\PwTc{} can be lowered efficiently to any architecture supported by \cXI{},
but inherits the top-level axioms of \rcXI{}, compromising compositionality.
\PwTmca{} is as a compositional as a model of concurrent imperative
programming can be, but is limited to \mca{} architectures for optimal
lowering.  It would be interesting to explore the middle ground to find a
fully compositional model that supports optimal lowering to all modern
architectures.

% In a similar vein, \PwT{} does not validate read-introduction, whereas
% speculative operational semantics do.  Recent work shows a tension between
% read-introduction and compositional reasoning for temporal safety properties
% (see \textsection\ref{sec:related}).  Nonetheless, read introduction is
% ubiquitous in some compilers.  It would be interesting to know if there is a
% performance penalty for banning read-introduction.

@@

Supplementary material for this paper available at
\url{https://weakmemory.github.io/pwt}.

@@

James Riely was supported by the National Science Foundation under
grant No.~CCR-1617175.

@@

Anton Podkopaev was supported by JetBrains Research.

### 2021-10-28 paper/related.tex
removing todos. Integrated consistency and causal strengtheining.

There are few prior studies of relaxed memory that include sequential
composition and/or precise calculation of semantic dependencies.  No prior
work gives the semantics of sequential composition in direct style.
\citet{DBLP:conf/esop/PaviottiCPWOB20} defined \MRD{}, which calculates
dependencies using event structures rather than logic.  This strategy is
brittler than ours, leading to false positives (\textsection\ref{sec:lir}).
\citet{DBLP:journals/pacmpl/JagadeesanJR20} defined \PwP{}, using logical
entailment to define dependency.  Although \PwT{} is based on \PwP{}, there
are many differences.  Two of these are motivated by requirements unique to
\PwT{}: First, \PwP{} uses substitution to define read actions; \PwT{}
Skolemizes in order to satisfy disjunction closure for predicate transformers
(\textsection\ref{sec:ex:assoc}).  Second, \PwP{} requires \emph{consistency}
and \emph{causal strengthening}; \PwT{} drops these in order to satisfy
associativity (\reflem{lem:monoid}).  Other differences are stylistic: For
example, we use termination \emph{conditions} rather than termination
\emph{actions}---our formulation fixes an error in
\citeauthor{DBLP:journals/pacmpl/JagadeesanJR20}'s definition of parallel
composition.  We also fix an error in their treatment of redundant read
elimination (\textsection\ref{sec:semreg}).

@@

% \todo{We agree that a new memory model needs to be positioned against existing
% models.  The usual result here is a compilation correctness to hardware
% memory models.  For PwT-MCA, we address this by showing compilation result
% for Armv8 model (§5).
% %
% Comparing software models, however, is unsatisfying: they are all
% incomparable, i.e., there are examples which are allowed by one/disallowed by
% the other and vice versa.
% %
% Morally, our model sits between the strong models (exemplified by RC11
% [Lahav-al:PLDI17]) and the speculative models (exemplified by the promising
% semantics [Kang-al:POPL17]).  As we argue in §1:
% %
% - The strong models require too much synchronization.
% %
% - The speculative models allow thin air behaviors.
% %
% Looking at the details, however, PwT-MCA is incomparable to both RC11 and
% promising semantics.  RC11 allows non-MCA behaviors that PwT-MCA disallows.
% PwT-MCA has a weaker notion of coherence than the promising semantics.
% %
% Some differences reflect an attempt to fix a bug.  For example, Weakestmo
% [Chakraborty-Vafeiadis:POPL19] purposefully disallows thin-air-like
% behaviors of the promising semantics.
% %
% Other differences reflect a different balance between allowed optimizations
% and reasoning principles.  There are fundamental conflicts, for example:
% %
% - between Common Subexpression Elimination (CSE) and read-read coherence
%   (§D.1)
% %
% - between if-introduction (aka, case analysis, if-closure) and java-style
%   final-field semantics.  (If-introduction requires that address dependencies
%   and control dependencies are the same.  Final-fields require that they be
%   different.)
% }
\citet{DBLP:journals/pacmpl/JagadeesanJR20} argue that the speculative models
allow too many executions, resulting in a failure of temporal reasoning and
potentially jeopardizing type safety and other security properties.  In a
similar vein, \citet{promising-ldrf} argue that local \drf{} guarantees are
violated when read-introduction is followed by if-introduction, branching on
the read just introduced.  These optimizations are validated by the
speculative models---\citeauthor{promising-ldrf} manage to avoid the problem
by adopting a sub-optimal lowering for \RMW{}s.  \PwT{} does not suffer from
this problem, since \PwT{} does not validate read-introduction.  Nonetheless,
read introduction is ubiquitous in some compilers
\cite{DBLP:conf/pldi/LeeKSHDMRL17}.  There appears to be a genuine tension
between temporal reasoning, as supported by \PwT{}, and read-introduction, as
supported by the speculative models.

>>>>

There are few prior studies of relaxed memory that include sequential
composition and/or precise calculation of semantic dependencies.  No prior
work gives the semantics of sequential composition in direct style.
\citet{DBLP:conf/esop/PaviottiCPWOB20} defined \MRD{}, which calculates
dependencies using event structures rather than logic.  This strategy is
brittler than ours, leading to false positives (\textsection\ref{sec:lir}).
\citet{DBLP:journals/pacmpl/JagadeesanJR20} defined \PwP{}, using logical
entailment to define dependency.  Although \PwT{} is based on \PwP{}, there
are many differences.
Some of these are motivated by requirements unique to
\PwT{} (see \textsection\ref{sec:ex:assoc}).
% Two of these are motivated by requirements unique to
% \PwT{}: First, \PwP{} uses substitution to define read actions; \PwT{}
% Skolemizes in order to satisfy disjunction closure for predicate transformers
% (\textsection\ref{sec:ex:assoc}).  Second, \PwP{} requires \emph{consistency}
% and \emph{causal strengthening}; \PwT{} drops these in order to satisfy
% associativity (\reflem{lem:monoid}).
Other differences are stylistic: For
example, we use termination \emph{conditions} rather than termination
\emph{actions}---our formulation fixes an error in
\citeauthor{DBLP:journals/pacmpl/JagadeesanJR20}'s definition of parallel
composition.  We also fix an error in their treatment of redundant read
elimination (\textsection\ref{sec:semreg}).

@@

% \todo{We agree that a new memory model needs to be positioned against existing
% models.  The usual result here is a compilation correctness to hardware
% memory models.  For PwT-MCA, we address this by showing compilation result
% for Armv8 model (§5).
% %
% Comparing software models, however, is unsatisfying: they are all
% incomparable, i.e., there are examples which are allowed by one/disallowed by
% the other and vice versa.
% %
% Morally, our model sits between the strong models (exemplified by RC11
% [Lahav-al:PLDI17]) and the speculative models (exemplified by the promising
% semantics [Kang-al:POPL17]).  As we argue in §1:
% %
% - The strong models require too much synchronization.
% %
% - The speculative models allow thin air behaviors.
% %
% Looking at the details, however, PwT-MCA is incomparable to both RC11 and
% promising semantics.  RC11 allows non-MCA behaviors that PwT-MCA disallows.
% PwT-MCA has a weaker notion of coherence than the promising semantics.
% %
% Some differences reflect an attempt to fix a bug.  For example, Weakestmo
% [Chakraborty-Vafeiadis:POPL19] purposefully disallows thin-air-like
% behaviors of the promising semantics.
% %
% Other differences reflect a different balance between allowed optimizations
% and reasoning principles.  There are fundamental conflicts, for example:
% %
% - between Common Subexpression Elimination (CSE) and read-read coherence
%   (§D.1)
% %
% - between if-introduction (aka, case analysis, if-closure) and java-style
%   final-field semantics.  (If-introduction requires that address dependencies
%   and control dependencies are the same.  Final-fields require that they be
%   different.)
% }
\citet{DBLP:journals/pacmpl/JagadeesanJR20} argue that the speculative models
allow too many executions, resulting in a failure of temporal reasoning and
potentially jeopardizing type safety and other security properties.  In a
similar vein, \citet{promising-ldrf} argue that local \drf{} guarantees are
violated when read-introduction is followed by if-introduction, branching on
the read just introduced.  These optimizations are validated by the
speculative models---\citeauthor{promising-ldrf} manage to avoid the problem
by adopting a sub-optimal lowering for \RMW{}s.  \PwT{} does not suffer from
this problem, since \PwT{} does not validate read-introduction.  Nonetheless,
read-introduction is ubiquitous in some compilers
\cite{DBLP:conf/pldi/LeeKSHDMRL17}.  There appears to be a genuine tension
between temporal reasoning, as supported by \PwT{}, and read-introduction, as
supported by the speculative models.

### 2021-10-28 paper/results.tex
removing todos. Integrated consistency and causal strengtheining.

\PwP{} %\cite{DBLP:journals/pacmpl/JagadeesanJR20}
is a novel memory model,
intended to serve as a semantic basis for a Java-like language, where all
access is safe.  \PwTmca{} generalizes \PwP{}, making several small but
significant changes.  As a result, we have had to re-prove most of the
theorems from \PwP{}.  The proofs can be found in \cite{appendix}.

% In \textsection\ref{sec:arm},
We show that \PwTmca{1} supports the optimal
lowering of relaxed accesses to \armeight{} and that \PwTmca{2} supports the
optimal lowering of \emph{all} accesses to \armeight{}.  The proofs are based
on two recent characterizations of \armeight{} \cite{armed}.  For \PwTmca{1},
we use \emph{External Global Consistency}.  For \PwTmca{2}, we use
\emph{External Consistency}.

%In \textsection\ref{sec:sc},
We prove sequential consistency for
local-data-race-free programs.  The proof uses \emph{program order}, which we
construct for \cXI{} in \textsection\ref{sec:c11}.  The same construction
works for \PwTmca{}.  (This proof assumes there are no \RMW{} operations.)

The semantics validates many peephole optimizations, such as the standard reorderings on relaxed access:
\begin{align*}
  %\taglabel{RR}
  \sembase{\PR{\aLoc}{\aReg} \SEMI \PR{\bLoc}{\bReg}} &=
  \sembase{\PR{\bLoc}{\bReg}\SEMI \PR{\aLoc}{\aReg}} &&\text{if } \aReg\neq\bReg
  \\
  %\taglabel{WW}
  \sembase{\aLoc \GETS \aExp \SEMI \bLoc  \GETS \bExp} &=
  \sembase{\bLoc  \GETS \bExp\SEMI \aLoc \GETS \aExp} &&\text{if } \aLoc\neq\bLoc
  \\
  %\taglabel{RW}
  \sembase{\aLoc \GETS \aExp  \SEMI \PR{\bLoc}{\bReg}} &=
  \sembase{\PR{\bLoc}{\bReg} \SEMI\aLoc \GETS \aExp} &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
\end{align*}
% \ref{WW}, \ref{RW} and \ref{RR} require that two sides of the semicolon
% have disjoint ids; for example, \ref{RW} requires $\disjoint{{\free(\aReg
% \GETS \aLoc)}}{{\free(\bLoc \GETS \bExp)}}$. 
% \ref{RR} requires either $\aReg\neq\bReg$ or
% $\aLoc=\bLoc$.  \ref{WW} and \ref{RW} require that two sides of the
% semicolon have disjoint ids; for example, \ref{RW} requires
% $\disjoint{{\free(\PR{\aLoc}{\aReg})}}{{\free(\bLoc \GETS \bExp)}}$.
%\ref{5} imposes no order between events in \ref{RR}--\ref{RW}.  %Note that \ref{RR} allows aliasing.
Here $\free(\aCmd)$ is the set of locations and registers that occur in $\aCmd$.
Using augmentation closure, the semantics also validates roach-motel reorderings \cite{SevcikThesis}.  For
example, on read/write pairs:
\begin{align*}
  % \tag{\textsc{roach1}}\label{AcqW}
  \sembase{x^\amode \GETS \aExp \SEMI\PR{y}{\bReg}} &\supseteq
  \sembase{\PR{y}{\bReg}  \SEMI x^\amode\GETS \aExp} 
  &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
  \\
  % \tag{\textsc{roach2}}\label{RelW}
  \sembase{x \GETS \aExp \SEMI\PR[\amode]{y}{\bReg}} &\supseteq
  \sembase{\PR[\amode]{y}{\bReg}  \SEMI x\GETS \aExp} 
  &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
\end{align*}

Notably, the semantics does \emph{not} validate read introduction.  When
combined with case analysis (\textsection\ref{sec:semca}), read introduction
can break temporal reasoning.  This combination is allowed by speculative
operational models.  See \textsection\ref{sec:related} for a discussion.

@@

>>>>

@@

% We \PwP{} %\cite{DBLP:journals/pacmpl/JagadeesanJR20}
% is a novel memory model, intended to serve as a semantic basis for a safe
% languages such as Java and Javascript.  \PwTmca{} generalizes \PwP{}, making
% several small but significant changes.  As a result, we have had to re-prove
% most of the theorems from \PwP{}.  These can be found in proofs can be found
% in \refAppendices{}.

In \refAppendix{sec:arm}, we show that \PwTmca{1} supports the optimal
lowering of relaxed accesses to \armeight{} and that \PwTmca{2} supports the
optimal lowering of \emph{all} accesses to \armeight{}.  The proofs are based
on two recent characterizations of \armeight{} \cite{armed}.  For \PwTmca{1},
we use \emph{External Global Consistency}.  For \PwTmca{2}, we use
\emph{External Consistency}.

In \refAppendix{sec:sc}, we also sketch a proof of sequential consistency for
local-data-race-free programs.  The proof uses \emph{program order}, which we
construct for \cXI{} in \textsection\ref{sec:c11}.  The same construction
works for \PwTmca{}.  (This proof sketch assumes there are no \RMW{} operations.)

The semantics validates many peephole optimizations, such as reorderings on relaxed access:
\begin{align*}
  %\taglabel{RR}
  \sembase{\PR{\aLoc}{\aReg} \SEMI \PR{\bLoc}{\bReg}} &=
  \sembase{\PR{\bLoc}{\bReg}\SEMI \PR{\aLoc}{\aReg}} &&\text{if } \aReg\neq\bReg
  \\
  %\taglabel{WW}
  \sembase{\aLoc \GETS \aExp \SEMI \bLoc  \GETS \bExp} &=
  \sembase{\bLoc  \GETS \bExp\SEMI \aLoc \GETS \aExp} &&\text{if } \aLoc\neq\bLoc
  \\
  %\taglabel{RW}
  \sembase{\aLoc \GETS \aExp  \SEMI \PR{\bLoc}{\bReg}} &=
  \sembase{\PR{\bLoc}{\bReg} \SEMI\aLoc \GETS \aExp} &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
\end{align*}
% \ref{WW}, \ref{RW} and \ref{RR} require that two sides of the semicolon
% have disjoint ids; for example, \ref{RW} requires $\disjoint{{\free(\aReg
% \GETS \aLoc)}}{{\free(\bLoc \GETS \bExp)}}$. 
% \ref{RR} requires either $\aReg\neq\bReg$ or
% $\aLoc=\bLoc$.  \ref{WW} and \ref{RW} require that two sides of the
% semicolon have disjoint ids; for example, \ref{RW} requires
% $\disjoint{{\free(\PR{\aLoc}{\aReg})}}{{\free(\bLoc \GETS \bExp)}}$.
%\ref{5} imposes no order between events in \ref{RR}--\ref{RW}.  %Note that \ref{RR} allows aliasing.
Here $\free(\aExp)$ is the set of locations and registers that occur in $\aExp$.
Using augmentation closure, the semantics also validates roach-motel reorderings \cite{SevcikThesis}.  For
example, on read/write pairs:
\begin{align*}
  % \tag{\textsc{roach1}}\label{AcqW}
  \sembase{x^\amode \GETS \aExp \SEMI\PR{y}{\bReg}} &\supseteq
  \sembase{\PR{y}{\bReg}  \SEMI x^\amode\GETS \aExp} 
  &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
  \\
  % \tag{\textsc{roach2}}\label{RelW}
  \sembase{x \GETS \aExp \SEMI\PR[\amode]{y}{\bReg}} &\supseteq
  \sembase{\PR[\amode]{y}{\bReg}  \SEMI x\GETS \aExp} 
  &&\text{if }
  \aLoc\neq\bLoc \textand \bReg\not\in\free(\aExp)%\disjoint{{\free(\aLoc \GETS \aExp)}}{{\free(\PR{\bLoc}{\bReg})}}
\end{align*}
Notably, the semantics does \emph{not} validate read-introduction.  When
combined with if-introduction (\textsection\ref{sec:semca}), read-introduction
can break temporal reasoning.  This combination is allowed by speculative
operational models.  See \textsection\ref{sec:related} for a discussion.
